---
title: "ea_data"
author: "Alejandra Garcia Isaza"
date: "10/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)

```

# Loading the data
```{r}
ea_raw <- read_sav(here("nopublishlea", "LEA EA W1.2 FINAL.sav"))

view_df(ea_raw)

data.frame(table(ea_raw$E0001.1)) # no duplicated ids

```

Participants that did not complete the survey:

EE05 = 15% completed
EE11 = 44% completed
EM16 = 30% completed

E2001.1 is where items from wave 1 begin

E2001.2 is where items from wave 2 begin, but E103.2
E103.2_Text is where wave 2 variables begin.


# Intial cleaning
```{r}
ea_clean <- ea_raw %>% 
  arrange(E0001.1) %>%
  select(-2:-9, -11) # selecting out columns with metadata
```

```{r}
describe(ea_clean)

```


```{r}

view_df(ea_clean)

# in wave 1:
ea_clean %>%
  filter(StdyCond == 0) # 25 control cases

ea_clean %>%
  filter(StdyCond == 1) # 20 control cases
```

Next steps:

- Identify which codes were used for no response, N/A, etc. -> appears that it's only -99
- Identify scales using HM's notes
- use parent.Rmd to follow code to compute scales
- Learn how to run EFA in R
- search "HM" to id questions I have for Heather

# recoding missing values as N/A with function
```{r}
# vector with missing values in dataset
missing_vals <- c(-99)

# function that returns true if values in vector are equal to missing_vals. The function takes a vector x, and specified values of missing data
recode_missing <- function(x, missing_vals = c(-99)) {
  test <- x %in% missing_vals
  ifelse(test, NA, x)
}

# function that recodes missing values to NA. The function takes a dataframe with variables with missing data, and specified values of missing data
recode_missing_df <- function(df, missing_vals = c(-99)) {
  modify(df, ~recode_missing(.x, missing_vals)) # here uses the function created above
}

ea_clean_na <- recode_missing_df(ea_clean) # applying this function strips out labels
```

# reverse scoring 

```{r}
ea_clean_rv <- ea_clean_na %>%
  mutate(E2009.1 = likert_reverse(E2009.1, top = 4, bottom = 1),
         E2013.1 = likert_reverse(E2013.1, top = 4, bottom = 1),
         E3004.1 = likert_reverse(E3004.1, top = 5, bottom = 1),
         E3005.1 = likert_reverse(E3005.1, top = 5, bottom = 1),
         E3007.1 = likert_reverse(E3007.1, top = 5, bottom = 1),
         E3008.1 = likert_reverse(E3008.1, top = 5, bottom = 1),
         E4003.1 = likert_reverse(E4003.1, top = 5, bottom = 1),
         E4024.1 = likert_reverse(E4024.1, top = 4, bottom = 1),
         E4027.1 = likert_reverse(E4027.1, top = 4, bottom = 1))
```

# Common guidelines for evaluating Cronbach's Alpha

.00 to .69 = Poor
.70 to .79 = Fair 
.80 to .89 = Good 
.90 to .99 = Excellent/Strong

### SCALES ###

# 01 = healthy emerging adulthood 
Items 1 - 16 in section 2
E2001.1 - E2016.1
response scale: 1 - 4 (strongly disagree - strongly agree)
higher scores = better 

item 9 = 	E2009.1 -> reverse scored
item 10 = E2010_1.1, E2010_2.1, E2010_3.1 -> excluded
item 13 = E2013.1 -> reverse scored

- Run EFA? -> EVALUATE WITH WED. DEADLINE. IF NOT, FACE VALIDITY AND SIMPLE RELIABILITY. FIRST RUN ALPHA, IF STRONG, KEEP. 

```{r}

s01 <- ea_clean_rv %>%
  select(E2001.1:E2016.1) %>% 
  select(-E2010_1.1, -E2010_2.1, -E2010_3.1)
  data.frame() 

alpha(s01, check.keys=FALSE) # 0.70

# E2009.1 is negatively correlated wth total scale despite rev score. The item could have been misinterpreted, so I am dropping it. 

s01 <- ea_clean_rv %>%
  select(E2001.1:E2016.1) %>% 
  select(-E2009.1,-E2010_1.1, -E2010_2.1, -E2010_3.1)
  data.frame() 

alpha(s01, check.keys=FALSE) # 0.78 healthy adulting

```

# 02 = PSS-10 (stress scale)
Items 1 - 10 in section 3
E3001.1 - E3010.1
response scale: 0 - 4 (never - very often)
higher scores = more stress

item 4 = E3004.1 - reverse scored
item 5 = E3005.1 - reverse scored
item 7 = E3007.1 - reverse scored
item 8 = E3008.1 - reverse scored

```{r}

s02 <- ea_clean_rv %>%
  select(E3001.1:E3010.1) %>% 
  data.frame() 

alpha(s02, check.keys=FALSE) # 0.82 stress
```

#### EA-parent relationship scales ####
items in section 4

# 03 Closeness/warmth scale in Aquilino (1999) -> check with HM need
item 1 = Global relationship quality = E4001.1
item 2 = humor = E4002.1
item 3 = tension = E4003.1 (needs reverse score)
item 5 = Shared activities = E4005.1

response scales: 
item 1 = 0 - 10 (really bad - absoluetly perfect)
item 2 = 1 - 5 (strongly disagree - strongly agree)
item 3 = 1 - 5 (strongly disagree - strongly agree) - reverse scored
item 5 = 1 - 5 (not at all - more than once a week)
higher scores = better 

***not sure if I can include these items in one scale given different response scales*** probably need to normalize -> I need to learn how to do this

```{r}
s03 <- ea_clean_rv %>%
  select(E4001.1, E4002.1, E4003.1, E4005.1) %>%
  data.frame()

alpha(s03, check.keys=FALSE) # 0.71 warmth (fair)

###### I am not completely sure if alpha can be done with items with different response options and response scales (1-5 and 0-10) ######
```

# 04 cultural diffs & conflict -> NOT USING THIS ONE
items 24 - 28
E4024.1 - E4028.1
response scale: 1 - 4 (strongly disagree - strongly agree)
higher scores = less conflict

item 24 = E4024.1 - reverse scored
item 27 = E4027.1 - reverse scored

```{r}
# s04 <- ea_clean_rv %>%
#   select(E4024.1:E4028.1) %>%
#   data.frame()
# 
# alpha(s04, check.keys=FALSE) # 0.21 cultural conflict

######## really poor alpha, hence not using this scale ########

```

# 05 (EA-Parent rel. quality)
items 21, 22 -> can these also be trust?
E4021.1, E4022.1
response scale: 1 - 4 (strongly disagree - strongly agree)

items 31, 32, 34 -> these can be trust
E4031.1, E4032.1, E4034.1 -> INCLUDE E4033.1 AND CHECK ALPHA
response scale: 0 - 3 (not at all true - very true) but same values
higher scores = more trust

```{r}
s05 <- ea_clean_rv %>%
  select(E4021.1, E4022.1, E4031.1, E4032.1, E4033.1, E4034.1) %>% 
  data.frame() 

alpha(s05, check.keys=FALSE) # 0.82 parent - ea rel. quality (trust?)
```

# 06 language comfort -> check with HM which
items 48 and 49 in section 5
E5048_1.1 - E5049_5.1
response scale: 1 - 5 (Not at all comf. - Very comf.)
higher scores = more comfort with each language

- Which ones we want to include: SPA/ENG at home, school, work, friends, in general? -> JUST FOCUS ON IN GENERAL

```{r}
ea_clean_rv$E5048_5.1 # comfort in SPA in general

ea_clean_rv$E5049_5.1 # comfort in ENG in general

```

# 07 friend group & alcohol use -> check with HM EFA
items 1 - 27 in section 6
E6001.1 - E6027.1
response scale: 1 - 4 (strongly disagree - strongly agree)

items 1 - 10 higher scores = more risky drinking (tendency to binge drink)
items 11 - 23 higher scores = more harm reduction strategies
items 24-27 are not as clear to me. 

- Run EFA - one scale and reverse score items 11 and above? or simply create 2 scales (risky drinking & harm reduction)?
- check with HM need of scale with items 28 - 41 about friend group? 

START WITH 2 SCALES

```{r}
s07 <- ea_clean_rv %>%
  select(E6001.1:E6010.1) %>% 
  data.frame() 

alpha(s07, check.keys=FALSE) # 0.93 tendency to binge drink among friends
```

```{r}
s07_2 <- ea_clean_rv %>%
  select(E6011.1:E6023.1) %>% 
  data.frame() 

alpha(s07_2, check.keys=FALSE) # 0.93 harm reduction strategies
```

# 08 alcohol use (Monitoring the future)
items 1 - 15 in section 7 
START IN E7001.1 -  JUST INDIVIDUAL ITEMS

```{r}
moni_fut <- ea_clean_rv %>%
  select(E0001.1, ResponseId, StdyCond, E0003.1, LocationRC.1, E7001.1:E7015.1) %>% 
  data.frame() 
```


# Demographic variables
items 1 - 24 in section 10
E1001_1.1
- HM: need to include these in the dataset?
- a lot of items are mark all that apply -> HM: LEAVE THEM AS THEY ARE

```{r}
dems <- ea_clean %>%
  select(E5048_5.1, E5049_5.1, E0001.1, ResponseId, StdyCond, E0003.1, LocationRC.1, E1001_1.1:E1024_5.1_TEXT) %>% 
  data.frame() # I did not apply the N/A function to not strip out variable labels
```


***What is the end result of the dataset?*** -> ask HM

- New dataset with only ids, condition, state, created scales (for wave 1 and 2), and demo vars

- Share syntax OF HOW SCALES WERE CONSTRUCTED with HM

### copying variable names for composites for HM

# Healthy EAing (alpha w1 = 0.78)
HLTYEA.1 = (E2001.1, E2002.1, E2003.1, E2004.1, E2005.1, E2006.1, E2007.1, E2008.1, E2011.1, E2012.1, E2013.1, E2014.1, E2015.1, E2016.1)

HLTYEA.2 = (E2001.2, E2002.2, E2003.2, E2004.2, E2005.2, E2006.2, E2007.2, E2008.2, E2011.2, E2012.2, E2013.2, E2014.2, E2015.2, E2016.2)

# PSS (alpha w1 = 0.82)
PSS.1 = (E3001.1, E3002.1, E3003.1, E3004.1, E3005.1, E3006.1, E3007.1, E3008.1, E3009.1, E3010.1)
PSS.2 = (E3001.2, E3002.2, E3003.2, E3004.2, E3005.2, E3006.2, E3007.2, E3008.2, E3009.2, E3010.2)

# EA-Parent rel. quality (trust) (alpha w1 = 0.82) 
RELQ.1 = (E4021.1, E4022.1, E4031.1, E4032.1, E4033.1, E4034.1)
RELQ.2 = (E4021.2, E4022.2, E4031.2, E4032.2, E4033.2, E4034.2)

# binge drinking (alpha w1 = 0.93)
BINGE.1 = (E6001.1, E6002.1, E6003.1, E6004.1, E6005.1, E6006.1, E6007.1, E6008.1, E6009.1, E6010.1)
BINGE.2 = (E6001.2, E6002.2, E6003.2, E6004.2, E6005.2, E6006.2, E6007.2, E6008.2, E6009.2, E6010.2)

# harm reduction strategies (alpha w1 = 0.93)
HARMRED.1 = (E6011.1, E6012.1, E6013.1, E6014.1, E6015.1, E6016.1, E6017.1, E6018.1, E6019.1, E6020.1, E6021.1, E6022.1, E6023.1)
HARMRED.1 = (E6011.2, E6012.2, E6013.2, E6014.2, E6015.2, E6016.2, E6017.2, E6018.2, E6019.2, E6020.2, E6021.2, E6022.2, E6023.2)

### composite scales with acceptable consistency

```{r}
composite_scales <- ea_clean_rv %>%
  rowwise() %>% 
  mutate(HLTYEA.1 = mean(c(E2001.1, E2002.1, E2003.1, E2004.1, E2005.1, E2006.1, E2007.1, E2008.1, E2011.1, E2012.1, E2013.1, E2014.1, E2015.1, E2016.1), na.rm = TRUE),
         HLTYEA.2 = mean(c(E2001.2, E2002.2, E2003.2, E2004.2, E2005.2, E2006.2, E2007.2, E2008.2, E2011.2, E2012.2, E2013.2, E2014.2, E2015.2, E2016.2), na.rm = TRUE),
         PSS.1 = mean(c(E3001.1, E3002.1, E3003.1, E3004.1, E3005.1, E3006.1, E3007.1, E3008.1, E3009.1, E3010.1), na.rm = TRUE),
         PSS.2 = mean(c(E3001.2, E3002.2, E3003.2, E3004.2, E3005.2, E3006.2, E3007.2, E3008.2, E3009.2, E3010.2), na.rm = TRUE),
         RELQ.1 = mean(c(E4021.1, E4022.1, E4031.1, E4032.1, E4033.1, E4034.1), na.rm = TRUE),
         RELQ.2 = mean(c(E4021.2, E4022.2, E4031.2, E4032.2, E4033.2, E4034.2), na.rm = TRUE),
         BINGE.1 = mean(c(E6001.1, E6002.1, E6003.1, E6004.1, E6005.1, E6006.1, E6007.1, E6008.1, E6009.1, E6010.1), na.rm = TRUE),
         BINGE.2 = mean(c(E6001.2, E6002.2, E6003.2, E6004.2, E6005.2, E6006.2, E6007.2, E6008.2, E6009.2, E6010.2), na.rm = TRUE),
         HARMRED.1 = mean(c(E6011.1, E6012.1, E6013.1, E6014.1, E6015.1, E6016.1, E6017.1, E6018.1, E6019.1, E6020.1, E6021.1, E6022.1, E6023.1), na.rm = TRUE),
         HARMRED.2 = mean(c(E6011.2, E6012.2, E6013.2, E6014.2, E6015.2, E6016.2, E6017.2, E6018.2, E6019.2, E6020.2, E6021.2, E6022.2, E6023.2), na.rm = TRUE)) %>%
  select(E0001.1, ResponseId, StdyCond, E0003.1, LocationRC.1, HLTYEA.1, HLTYEA.2, PSS.1, PSS.2, RELQ.1, RELQ.2, BINGE.1, BINGE.2, HARMRED.1, HARMRED.2) # selecting only id variables and the new scales

```


```{r}

final_ea <- left_join(ea_clean, composite_scales)

view_df(final_ea)
```

# exporting dataset to SPSS
```{r}
final_ea %>%
  haven::write_sav(here("nopublishlea", "final_ea.sav")) 
```
