---
title: "ea_data"
author: "Alejandra Garcia Isaza"
date: "10/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)

```

# Loading the data
```{r}
ea_raw <- read_sav(here("nopublishlea", "LEA EA W1.2 FINAL.sav"))

view_df(ea_raw)

data.frame(table(ea_raw$E0001.1)) # no duplicated ids

```

Participants that did not complete the survey:

EE05 = 15% completed
EE11 = 44% completed
EM16 = 30% completed

E2001.1 is where items from wave 1 begin

E2001.2 is where items from wave 2 begin, but E103.2
E103.2_Text is where wave 2 variables begin.


# Intial cleaning
```{r}
ea_clean <- ea_raw %>% 
  arrange(E0001.1) %>%
  select(-2:-9, -11) # selecting out columns with metadata
```

```{r}
describe(ea_clean)
```


```{r}

view_df(ea_clean)

# in wave 1:
ea_clean %>%
  filter(StdyCond == 0) # 25 control cases

ea_clean %>%
  filter(StdyCond == 1) # 20 control cases
```

Next steps:

- Identify which codes were used for no response, N/A, etc. -> appears that it's only -99
- Identify scales using HM's notes
- use parent.Rmd to follow code to compute scales
- Learn how to run EFA in R
- search "HM" to id questions I have for Heather

### SCALES ###

# 01 = healthy emerging adulthood 
Items 1 - 16 in section 2
E2001.1 - E2016.1
response scale: 1 - 4 (strongly disagree - strongly agree)
higher scores = better 

item 9 = 	E2009.1 -> probably needs reverse scoring -> check with HM
item 10 = E2010_1.1, E2010_2.1, E2010_3.1 -> exclude (fill in blank)
item 13 = E2013.1 -> needs reverse scoring 

- Run EFA?


# 02 = PSS-10 (stress scale) -> check with HM reverse scoring items
Items 1 - 10 in section 3
E3001.1 - E3010.1
response scale: 0 - 4 (never - very often)
higher scores = more stress

item 4 = E3004.1 - needs reverse scoring
item 5 = E3005.1 - needs reverse scoring
item 7 = E3007.1 - needs reverse scoring
item 8 = E3008.1 - needs reverse scoring


#### EA-parent relationship scales ####
items in section 4

# 03 Closeness/warmth scale in Aquilino (1999) -> check with HM need
item 1 = Global relationship quality = E4001.1
item 2 = humor = E4002.1
item 3 = Tension (reverse coded) = E4003.1
item 5 = Shared activities = E4005.1

response scales: 
item 1 = 0 - 10 (really bad - absoluetly perfect)
items 2 and 3 = 1 - 5 (strongly disagree - strongly agree)
item 5 = 1 - 5 (not at all - more than once a week)
higher scores = better 

***not sure if I can include these items in one scale given different response scales*** probably need to normalize -> I need to learn how to do this

# 04 cultural diffs & conflict -> check with HM need
items 24 - 28
E4024.1 - E4028.1
response scale: 1 - 4 (strongly disagree - strongly agree)
higher scores = less conflict

item 24 = E4024.1 - needs reverse scoring
item 27 = E4027.1 - needs reverse scoring

# 05 (?) -> check items with HM to come with a name
items 21, 22 -> can these also be trust?
E4021.1, E4022.1
response scale: 1 - 4 (strongly disagree - strongly agree)

items 31, 32, 34 -> these can be trust
E4031.1, E4032.1, E4034.1
response scale: 0 - 3 (not at all true - very true)
higher scores = more (?)

# 06 language comfort -> check with HM which
items 48 and 49 in section 5
E5048_1.1 - E5049_5.1
response scale: 1 - 5 (Not at all comf. - Very comf.)
higher scores = more comfort with each language

- Which ones we want to include: SPA/ENG at home, school, work, friends, in general?
- Or make a composite of all?

# 07 friend group & alcohol use -> check with HM EFA
items 1 - 27 in section 6
E6001.1 - E6027.1
response scale: 1 - 4 (strongly disagree - strongly agree)

items 1 - 10 higher scores = more risky drinking 
items 11 - 22 higher scores = more harm reduction strategies
items 24-27 are not as clear to me. 

- Run EFA - one scale and reverse score items 11 and above? or simply create 2 scales (risky drinking & harm reduction)?
- check with HM need of scale with items 28 - 41 about friend group? 

# 08 alcohol use (Monitoring the future)
items 1 - 15 in section 7 

- HM: choose items or run EFA?

# Demographics variables
items 1 - 24 in section 10
E1001_1.1
- HM: need to include these in the dataset?
- a lot of items are mark all that apply -> HM: what to do?


***What is the end result of the dataset?*** -> ask HM

Attach the scales I create at the end of the original dataset or create a new dataset with only ids, condition, state, created scales (for wave 1 and 2, right?), and demo vars?


