---
title: "parent_data"
author: "Alejandra Garcia Isaza"
date: "11/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
```


# Loading the data
```{r}
pa_raw <- read_sav(here("nopublishlea", "LEA Parent W1.2 FINAL.sav"))

# data.frame(table(pa_raw$P0001.1)) # no duplicated ids

# pa_raw %>%
#   filter(Finished == 0)
```

Participants that did not complete the survey:

PM05 = 87% completed

P2001.1 is where items from wave 1 begin

P2001.2 is where items from wave 2 begin, but P1003.2
P1003.2_TEXT is where wave 2 variables begin.

# Intial cleaning
```{r, echo=T, results='hide'}
pa_clean <- pa_raw %>% 
  arrange(P0001.1) %>%
  select(-2:-9, -11) %>% # selecting out columns with metadata
  filter(P0001.1 != "XXXX") # removing data with odd id
```

```{r, echo=T, results='hide'}

# in wave 1:
pa_clean %>%
  filter(E0002.1 == 0) # 19 control cases

pa_clean %>%
  filter(E0002.1 == 1) # 19 intervention cases
```

<!-- between P2007.1 and P2009.1, the values get mixed up. The scale was 1,2,3,4,5 and what was coded on Qualtrics was 1,2,4,6,7 (needs fixing, but later) --> 

between P2021.1 and P2033.1 (& w2: P2021.2 and P2033.2) the values get mixed up. The scale was 1,2,3,4 and what was coded on Qualtrics was 1,2,4,5 --> needs fixing NOW

```{r}
fix_1 <- pa_clean %>%
  select(P0001.1, P2021.1:P2033.1 | P2021.2:P2033.2)
```


```{r}
# recoding incorrect values as N/A with function

# vector with incorrect values in dataset
vals <- c(4)

# function that returns true if values in vector are equal to vals. The function takes a vector x, and specified values of incorrect data
recode_incorrect <- function(x, vals = c(4)) {
  test <- x %in% vals
  ifelse(test, 3, x)
}

# function that recodes incorrect values to 3. The function takes a dataframe with variables with incorrect data, and specified values of incorrect data
recode_incorrect_df <- function(df, vals = c(4)) {
  modify(df, ~recode_incorrect(.x, vals)) # here uses the function created above
}

fix_1 <- recode_incorrect_df(fix_1) # applying this function strips out labels
```

```{r}
# recoding incorrect values as N/A with function

# vector with incorrect values in dataset
vals <- c(5)

# function that returns true if values in vector are equal to vals. The function takes a vector x, and specified values of incorrect data
recode_incorrect <- function(x, vals = c(5)) {
  test <- x %in% vals
  ifelse(test, 4, x)
}

# function that recodes incorrect values to 3. The function takes a dataframe with variables with incorrect data, and specified values of incorrect data
recode_incorrect_df <- function(df, vals = c(5)) {
  modify(df, ~recode_incorrect(.x, vals)) # here uses the function created above
}

fix_1.2 <- recode_incorrect_df(fix_1) # applying this function strips out labels
```

between P2034.1 and P2037.1 (& w2: P2034.2 and P2037.2) the values get mixed up. The scale was 1,2,3,4 and what was coded on Qualtrics was 1,2,6,7 --> needs fixing NOW

```{r}
fix_2 <- pa_clean %>%
  select(P0001.1, P2034.1:P2037.1 | P2034.2:P2037.2)
```


```{r}
# recoding incorrect values as N/A with function

# vector with incorrect values in dataset
vals <- c(6)

# function that returns true if values in vector are equal to vals. The function takes a vector x, and specified values of incorrect data
recode_incorrect <- function(x, vals = c(6)) {
  test <- x %in% vals
  ifelse(test, 3, x)
}

# function that recodes incorrect values to 3. The function takes a dataframe with variables with incorrect data, and specified values of incorrect data
recode_incorrect_df <- function(df, vals = c(6)) {
  modify(df, ~recode_incorrect(.x, vals)) # here uses the function created above
}

fix_2 <- recode_incorrect_df(fix_2) # applying this function strips out labels
```


```{r}
# recoding incorrect values as N/A with function

# vector with incorrect values in dataset
vals <- c(7)

# function that returns true if values in vector are equal to vals. The function takes a vector x, and specified values of incorrect data
recode_incorrect <- function(x, vals = c(7)) {
  test <- x %in% vals
  ifelse(test, 4, x)
}

# function that recodes incorrect values to 3. The function takes a dataframe with variables with incorrect data, and specified values of incorrect data
recode_incorrect_df <- function(df, vals = c(7)) {
  modify(df, ~recode_incorrect(.x, vals)) # here uses the function created above
}

fix_2.2 <- recode_incorrect_df(fix_2) # applying this function strips out labels
```

P2038.1 and P2062.1 (& w2: P2038.2 and P2062.2)the values get mixed up. The scale was 1,2,3,4 and what was coded on Qualtrics was 1,3,5,7 --> needs fixing NOW

```{r}
fix_3 <- pa_clean %>%
  select(P0001.1, P2038.1:P2062.1 | P2038.2:P2062.2)
```


```{r}
# recoding incorrect values as N/A with function

# vector with incorrect values in dataset
vals <- c(3)

# function that returns true if values in vector are equal to vals. The function takes a vector x, and specified values of incorrect data
recode_incorrect <- function(x, vals = c(3)) {
  test <- x %in% vals
  ifelse(test, 2, x)
}

# function that recodes incorrect values to 3. The function takes a dataframe with variables with incorrect data, and specified values of incorrect data
recode_incorrect_df <- function(df, vals = c(3)) {
  modify(df, ~recode_incorrect(.x, vals)) # here uses the function created above
}

fix_3 <- recode_incorrect_df(fix_3) # applying this function strips out labels
```

```{r}
# recoding incorrect values as N/A with function

# vector with incorrect values in dataset
vals <- c(5)

# function that returns true if values in vector are equal to vals. The function takes a vector x, and specified values of incorrect data
recode_incorrect <- function(x, vals = c(5)) {
  test <- x %in% vals
  ifelse(test, 3, x)
}

# function that recodes incorrect values to 3. The function takes a dataframe with variables with incorrect data, and specified values of incorrect data
recode_incorrect_df <- function(df, vals = c(5)) {
  modify(df, ~recode_incorrect(.x, vals)) # here uses the function created above
}

fix_3.2 <- recode_incorrect_df(fix_3) # applying this function strips out labels
```

```{r}
# recoding incorrect values as N/A with function

# vector with incorrect values in dataset
vals <- c(7)

# function that returns true if values in vector are equal to vals. The function takes a vector x, and specified values of incorrect data
recode_incorrect <- function(x, vals = c(7)) {
  test <- x %in% vals
  ifelse(test, 4, x)
}

# function that recodes incorrect values to 3. The function takes a dataframe with variables with incorrect data, and specified values of incorrect data
recode_incorrect_df <- function(df, vals = c(7)) {
  modify(df, ~recode_incorrect(.x, vals)) # here uses the function created above
}

fix_3.3 <- recode_incorrect_df(fix_3.2) # applying this function strips out labels
```

```{r}
d_fixed_1 <- left_join(fix_1.2, fix_2.2)

d_fixed_2 <- left_join(d_fixed_1, fix_3.3) 

other_vars <- pa_clean %>%
  select(P0001.1, P4013_5.1, P4014_5.1, P4013_5.2, P4014_5.2, P3001.1:P3010.1, P3001.2:P3010.2, P7001.1:P70012.1, P7001.2:P70012.2)

d_fixed_3 <- left_join(d_fixed_2, other_vars)

```

# recoding missing values as N/A with function
```{r}
# vector with missing values in dataset
missing_vals <- c(-99)

# function that returns true if values in vector are equal to missing_vals. The function takes a vector x, and specified values of missing data
recode_missing <- function(x, missing_vals = c(-99)) {
  test <- x %in% missing_vals
  ifelse(test, NA, x)
}

# function that recodes missing values to NA. The function takes a dataframe with variables with missing data, and specified values of missing data
recode_missing_df <- function(df, missing_vals = c(-99)) {
  modify(df, ~recode_missing(.x, missing_vals)) # here uses the function created above
}

d_fixed_na <- recode_missing_df(d_fixed_3) # applying this function strips out labels

#view_df(d_fixed_na) # it worked, all items are in 1-4 scales and no -99
```

# reverse scoring 

```{r}
d_fixed_na_rv <- d_fixed_na %>%
  mutate(P2059.1 = likert_reverse(P2059.1, top = 4, bottom = 1),
         P2061.1 = likert_reverse(P2061.1, top = 4, bottom = 1),
         P3004.1 = likert_reverse(P3004.1, top = 5, bottom = 1),
         P3005.1 = likert_reverse(P3005.1, top = 5, bottom = 1),
         P3007.1 = likert_reverse(P3007.1, top = 5, bottom = 1),
         P3008.1 = likert_reverse(P3008.1, top = 5, bottom = 1),
         P70010.1 = likert_reverse(P70010.1, top = 4, bottom = 1),
         P70011.1 = likert_reverse(P70011.1, top = 4, bottom = 1))
```

# Common guidelines for evaluating Cronbach's Alpha

.00 to .69 = Poor
.70 to .79 = Fair 
.80 to .89 = Good 
.90 to .99 = Excellent/Strong

### SCALES ###

# Mentor parenting scales
Items 1 - 62 in section 2
P2021.1 - P2042.1
response scale: 1 - 4 (strongly disagree - strongly agree)

# confidence in one's parenting
higher scores = better
```{r}
s01 <- d_fixed_na_rv %>%
  select(P2021.1, P2025.1, P2027.1, P2028.1, P2029.1, P2032.1, P2033.1) %>%
  data.frame()

alpha(s01, check.keys=FALSE) # 0.81 (good)

```

# supporting EA
higher scores = better
```{r}
# s02 <- d_fixed_na_rv %>%
#   select(P2023.1, P2024.1, P2026.1, P2034.1, P2035.1, P2036.1, P2037.1, P2044.1, P2046.1) %>%
#   data.frame()
# 
# alpha(s02, check.keys=FALSE) # 0.77

# removing P2044.1 because it was negatively related to the scale, even though it didn't need to be reverse scored --> the item was a little "tricky". 

s02 <- d_fixed_na_rv %>%
  select(P2023.1, P2024.1, P2026.1, P2034.1, P2035.1, P2036.1, P2037.1, P2046.1) %>%
  data.frame()

alpha(s02, check.keys=FALSE) # 0.82 (good)
```

# treating EA as a child
higher scores = worse 

```{r}
s03 <- d_fixed_na_rv %>%
  select(P2038.1, P2040.1, P2045.1, P2055.1) %>%
  data.frame()

alpha(s03, check.keys=FALSE) # 0.54 (poor)

# Not using this one, even if taking out P2045.1 = 0.67 (poor) #
```

# supporting decision-making
higher scores = better
```{r}
s04 <- d_fixed_na_rv %>%
  select(P2039.1, P2053.1, P2054.1, P2056.1) %>%
  data.frame()

alpha(s04, check.keys=FALSE) # 0.46 (poor)

# Not using this one, even if taking out P2039.1 = 0.56(poor) #

```

# parenting & EA risky behavior 
higher scores = better
```{r}
s05 <- d_fixed_na_rv %>%
  select(P2058.1, P2059.1, P2060.1, P2061.1, P2062.1) %>%
  data.frame()

alpha(s05, check.keys=FALSE) #0.51 (poor)

# Not using this one, even if taking out P2059.1 = 0.69(poor) #
```

# 02 = PSS-10 (stress scale)
Items 1 - 10 in section 3
P3001.1 - P3010.1
response scale: 0 - 4 (never - very often) --> but coded in qualtrics as 1 - 5 
higher scores = more stress

item 4 = E3004.1 - reverse scored
item 5 = E3005.1 - reverse scored
item 7 = E3007.1 - reverse scored
item 8 = E3008.1 - reverse scored

```{r}

s06 <- d_fixed_na_rv %>%
  select(P3001.1:P3010.1) %>% 
  data.frame() 

alpha(s06, check.keys=FALSE) # 0.83 stress
```

# SPA general
P4013_5.1 --> response scales OK

# ENG general
P4014_5.1 --> response scales OK


# trust in EAs' decisions 
P7001.1 - P70012.1
response scale: 0 - 3 (not at all true - very true) --> but coded in qualtrics as 1 - 4

```{r}

s06 <- d_fixed_na_rv %>%
  select(P7007.1, P7008.1, P7009.1, P70010.1, P70011.1, P70012.1) %>% 
  data.frame() 

alpha(s06, check.keys=FALSE) # 0.77 (fair)
```

### composite scales with acceptable consistency

```{r}
composite_scales <- d_fixed_na_rv %>%
  rowwise() %>% 
  mutate(CONF.1 = mean(c(P2021.1, P2025.1, P2027.1, P2028.1, P2029.1, P2032.1, P2033.1), na.rm = TRUE),
         CONF.2 = mean(c(P2021.2, P2025.2, P2027.2, P2028.2, P2029.2, P2032.2, P2033.2), na.rm = TRUE),
         SUPP.1 = mean(c(P2023.1, P2024.1, P2026.1, P2034.1, P2035.1, P2036.1, P2037.1, P2046.1), na.rm = TRUE),
         SUPP.2 = mean(c(P2023.2, P2024.2, P2026.2, P2034.2, P2035.2, P2036.2, P2037.2, P2046.2), na.rm = TRUE),
         PSS.1 = mean(c(P3001.1, P3002.1, P3003.1, P3004.1, P3005.1, P3006.1, P3007.1, P3008.1, P3009.1, P3010.1), na.rm = TRUE),
         PSS.2 = mean(c(P3001.2, P3002.2, P3003.2, P3004.2, P3005.2, P3006.2, P3007.2, P3008.2, P3009.2, P3010.2), na.rm = TRUE),
         TRU.1 = mean(c(P7007.1, P7008.1, P7009.1, P70010.1, P70011.1, P70012.1), na.rm = TRUE),
         TRU.2 = mean(c(P7007.1, P7008.1, P7009.1, P70010.1, P70011.1, P70012.1), na.rm = TRUE)) 

```

```{r}
id_vars <- pa_clean %>%
  select(P0001.1, E0002.1, E0003.1, P1003.1, P1003.1_TEXT, LocationRC.1)

reduced_pa <- left_join(id_vars, composite_scales)

```


```{r, message=FALSE}
view_df(pa_clean)
```

# exporting reduced dataset to SPSS (only variables that were cleaned)
```{r}
# reduced_pa %>%
#   haven::write_sav(here("nopublishlea", "reduced_pa.sav"))
```
